---
title: "Bellabeat Smart Device Analysis: Unlocking Growth Opportunities"
author: "Data Analysis Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 10, fig.height = 6)
```

# Executive Summary

This analysis examines FitBit smart device usage data from 30 users to understand activity patterns, sleep behavior, and feature utilization. The primary objective is to inform Bellabeat's marketing strategy and product development by identifying how consumers use non-Bellabeat smart devices.

Key findings reveal distinct user segments based on activity levels, significant correlations between activity and calorie burn, and varying feature adoption rates. These insights provide actionable opportunities for Bellabeat to enhance their product offerings and targeted marketing campaigns.

---

# 1. Ask Phase

## 1.1 Business Task

Bellabeat, a high-tech manufacturer of health-focused products for women, seeks to expand its market presence. This analysis addresses the fundamental question: **How do consumers use smart fitness devices, and how can these insights guide Bellabeat's marketing strategy?**

## 1.2 Stakeholders

- **Primary Stakeholders**: Urška Sršen (Cofounder & Chief Creative Officer), Sando Mur (Cofounder & Mathematician), Bellabeat Executive Team
- **Secondary Stakeholders**: Bellabeat Marketing Analytics Team

## 1.3 Key Questions

1. What are the trends in smart device usage?
2. How could these trends apply to Bellabeat customers?
3. How could these trends help influence Bellabeat's marketing strategy?
4. What features do users value most?
5. What usage patterns indicate opportunities for product development?

---

# 2. Prepare Phase

## 2.1 Data Source

The analysis utilizes FitBit Fitness Tracker Data made available through Mobius on [Kaggle](https://www.kaggle.com/datasets/arashnic/fitbit). The dataset contains personal fitness tracker data from thirty FitBit users who consented to the submission of their personal tracker data via Amazon Mechanical Turk between March 12, 2016 and May 12, 2016.

## 2.2 Data Description

- **Format**: 18 CSV files containing different granularities of tracking data
- **Storage Location**: `data/raw/` directory
- **Time Period**: March 12, 2016 - May 12, 2016 (2 months)
- **Users**: 30 participants
- **Tracking Range**: 4-31 days per user

### Dataset Overview

| Dataset | Granularity | Key Metrics |
|---------|-------------|-------------|
| Daily Activity | Daily | Steps, distance, calories, active minutes |
| Daily Calories | Daily | Calorie burn |
| Daily Intensities | Daily | Activity intensity levels |
| Daily Steps | Daily | Total steps |
| Sleep Day | Daily | Sleep duration, efficiency |
| Weight Log | Variable | Weight, BMI, manual/auto logs |
| Hourly Data | Hourly | Steps, calories, intensities |
| Minute Data | Per minute | Detailed activity, sleep, METs |
| Heart Rate | Per second | Continuous heart rate monitoring |

## 2.3 Data Limitations

**Critical Acknowledgments**:
- Small sample size (30 users) limits generalizability
- Dataset from 2016 may not reflect current trends
- No demographic information (age, gender, location)
- Only 2 months of data; cannot capture seasonal or long-term patterns
- Inconsistent feature usage across users
- Variable tracking duration (4-31 days per user)
- Not all users tracked all features (sleep: 80%, heart rate: 47%, weight: 27%)

---

# 3. Process Phase

## 3.1 Tools Used

```{r libraries}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(forcats)
library(vroom)
library(janitor)
library(here)
library(readr)
library(scales)
library(broom)
library(knitr)
```

## 3.2 Data Cleaning Process

The data cleaning pipeline (`scripts/01_data_cleaning.R`) performed the following operations:

### 3.2.1 Data Loading

```{r eval=FALSE}
# Load all 18 CSV files
file_paths <- list.files(
  here("data", "raw"),
  pattern = "\\.csv$",
  full.names = TRUE
)

bellabeat_file_data_list <- map(file_paths, ~ vroom(.x, progress = FALSE))
names(bellabeat_file_data_list) <- basename(file_paths)
```

### 3.2.2 Data Transformation

1. **Column Standardization**
   - Cleaned column names using `janitor::clean_names()`
   - Converted IDs to character type for consistency
   - Standardized datetime formats across all datasets

2. **Data Validation & Cleaning**
   - Removed duplicate records
   - Filtered invalid heart rates (>220 bpm or ≤0)
   - Removed days with total minutes >1440 (impossible values)
   - Removed whitespace and newline characters

3. **Feature Engineering**
   - Created `total_active_minutes` (sum of very/fairly/lightly active)
   - Calculated `active_ratio` (active minutes / total minutes)
   - Computed `sleep_efficiency` (time asleep / time in bed)
   - Flagged `non_wear` days (all activity metrics = 0)
   - Rounded distance measurements to 2 decimal places

```{r eval=FALSE}
# Example feature engineering
daily_activities <- daily_activities %>%
  mutate(
    total_active_minutes = very_active_minutes + 
                          fairly_active_minutes + 
                          lightly_active_minutes,
    active_ratio = total_active_minutes / total_minutes,
    non_wear = (very_active_minutes == 0 & 
                fairly_active_minutes == 0 & 
                lightly_active_minutes == 0 & 
                total_steps == 0)
  ) %>%
  filter(total_minutes <= 1440)
```

### 3.2.3 Data Output

- **Cleaned Datasets**: 18 cleaned CSV files in `data/cleaned/`
- **Exploration Outputs**: Summary CSVs in `data/processed/`

---

# 4. Analyze Phase

## 4.1 Analysis Pipeline

### 4.1.1 Phase 1: Exploration (`scripts/02_exploration.R`)

**A. Data Coverage Analysis**

```{r eval=FALSE}
# User tracking consistency
user_tracking_days <- daily_activities %>%
  group_by(id) %>%
  summarize(
    days_tracked = n(),
    first_day = min(activity_day),
    last_day = max(activity_day)
  )

# Engagement classification
user_engagement <- user_tracking_days %>%
  mutate(
    engagement_level = case_when(
      days_tracked >= 25 ~ "Everyday User (25-31 days)",
      days_tracked >= 21 ~ "Heavy User (21-24 days)",
      days_tracked >= 11 ~ "Moderate User (11-20 days)",
      TRUE ~ "Light User (1-10 days)"
    )
  )
```

**Key Findings**:
- Tracking days range: 4-31 days per user
- Average tracking: ~24 days
- Engagement distribution across 4 levels

**B. Feature Utilization Analysis**

```{r eval=FALSE}
# Feature availability per user
feature_matrix <- tibble(id = all_users) %>%
  mutate(
    has_activity = id %in% daily_activities$id,
    has_sleep = id %in% sleep_day$id,
    has_weight = id %in% weight_log$id,
    has_heartrate = id %in% heartrate_seconds$id
  )
```

**Key Findings**:
- Activity tracking: 100% (all 30 users)
- Sleep tracking: ~80% (24 users)
- Heart rate monitoring: ~47% (14 users)
- Weight logging: ~27% (8 users)

**C. Basic Distributions**

```{r eval=FALSE}
# Activity distributions
activity_distributions <- daily_activities %>%
  summarize(
    steps_mean = mean(total_steps, na.rm = TRUE),
    steps_median = median(total_steps, na.rm = TRUE),
    calories_mean = mean(calories, na.rm = TRUE),
    sedentary_mean = mean(sedentary_minutes, na.rm = TRUE)
  )
```

**Key Findings**:
- Average daily steps: ~7,500 steps
- Average calories burned: ~2,300 calories
- Average sedentary time: ~50% of tracked minutes
- Average sleep: ~7 hours per night

### 4.1.2 Phase 2: Statistical Analysis (`scripts/03_analysis.R`)

**A. User Segmentation**

```{r eval=FALSE}
# Activity level classification
user_activity_segments <- daily_activities %>%
  group_by(id) %>%
  summarize(avg_daily_steps = mean(total_steps)) %>%
  mutate(
    activity_level = case_when(
      avg_daily_steps < 5000 ~ "Sedentary",
      avg_daily_steps < 7500 ~ "Low Active",
      avg_daily_steps < 10000 ~ "Somewhat Active",
      avg_daily_steps < 12500 ~ "Active",
      TRUE ~ "Highly Active"
    )
  )

# Sleep behavior classification
user_sleep_segments <- sleep_day %>%
  mutate(sleep_hours = total_minutes_asleep / 60) %>%
  group_by(id) %>%
  summarize(avg_sleep_hours = mean(sleep_hours)) %>%
  mutate(
    sleep_pattern = case_when(
      avg_sleep_hours < 6 ~ "Under-sleeper",
      avg_sleep_hours <= 8 ~ "Normal sleeper",
      TRUE ~ "Over-sleeper"
    )
  )
```

**Key Findings**:
- Activity levels range from Sedentary to Highly Active
- Most users fall into "Somewhat Active" or "Active" categories
- Majority of users are "Normal sleepers" (6-8 hours)

**B. Correlation Analysis**

```{r eval=FALSE}
# Key correlations
key_correlations <- tibble(
  relationship = c(
    "Steps → Calories",
    "Very Active Minutes → Calories",
    "Distance → Calories"
  ),
  correlation = c(
    cor(daily_activities$total_steps, daily_activities$calories),
    cor(daily_activities$very_active_minutes, daily_activities$calories),
    cor(daily_activities$total_distance, daily_activities$calories)
  )
)
```

**Key Findings**:
- Strong positive correlation: Steps → Calories (r > 0.7)
- Strong positive correlation: Very Active Minutes → Calories (r > 0.7)
- Moderate correlation: Sedentary Minutes → Calories (negative)
- Weak to moderate correlations between activity and sleep quality

**C. Temporal Patterns**

```{r eval=FALSE}
# Hourly activity summary
hourly_activity_summary <- hourly_steps %>%
  mutate(hour = hour(activity_hour)) %>%
  group_by(hour) %>%
  summarize(avg_steps = mean(step_total))

# Day of week patterns
daily_averages_by_dow <- daily_activities %>%
  mutate(day_of_week = wday(activity_day, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarize(avg_steps = mean(total_steps))
```

**Key Findings**:
- Peak activity hours: 12 PM - 7 PM
- Higher activity on weekdays vs weekends (commuting pattern)
- Consistent activity levels across the 2-month period
- No significant weekly drop-off in user engagement

**D. Sleep Analysis**

```{r eval=FALSE}
# Sleep efficiency calculation
sleep_analysis <- sleep_day %>%
  mutate(
    sleep_hours = total_minutes_asleep / 60,
    time_awake_in_bed = total_time_in_bed - total_minutes_asleep
  ) %>%
  summarize(
    avg_sleep_hours = mean(sleep_hours),
    avg_sleep_efficiency = mean(sleep_efficiency),
    avg_time_awake = mean(time_awake_in_bed)
  )
```

**Key Findings**:
- Average sleep duration: ~7 hours
- Average sleep efficiency: ~92%
- Average time awake in bed: ~39 minutes
- Sleep consistency varies significantly across users

## 4.2 Summary Statistics

```{r eval=FALSE, echo=FALSE}
# Load summary statistics
member_stats <- read_csv(here("data/processed/user_segments.csv"))
kable(head(member_stats), caption = "Sample User Segments")
```

---

# 5. Share Phase

## 5.1 Visualization Strategy

All visualizations were generated using `scripts/04_visualization.R` and stored in the `plots/` directory. The visualization approach focuses on:
- Clear communication of user behavior patterns
- Actionable insights for marketing and product development
- Consistent branding with custom `theme_bellabeat()`

## 5.2 Key Visualizations

### 5.2.1 User Overview Dashboard

**Users Per Dataset**
```{r eval=FALSE}
ggplot(users_per_dataset, aes(x = reorder(Dataset, Users), y = Users)) +
  geom_col(fill = "#1976D2", alpha = 0.8) +
  geom_text(aes(label = Users), hjust = -0.2) +
  coord_flip() +
  labs(title = "users per dataset",
       subtitle = "fitbit tracking coverage across data types") +
  theme_bellabeat()
```

**Insight**: Activity tracking is universal (100%), but sleep (80%), heart rate (47%), and weight (27%) show declining adoption.

**User Engagement Distribution**
```{r eval=FALSE}
ggplot(engagement_counts, aes(x = "", y = n, fill = engagement_level)) +
  geom_col(width = 1, color = "white") +
  coord_polar("y") +
  scale_fill_manual(values = colors_engagement) +
  labs(title = "user engagement distribution") +
  theme_void()
```

**Insight**: Users are fairly evenly distributed across engagement levels, with a slight majority being "Moderate" or "Heavy" users.

### 5.2.2 Activity Patterns

**Activity Heatmap**
```{r eval=FALSE}
ggplot(hourly_steps_data, 
       aes(x = hour, y = day_of_week, fill = avg_steps)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F57C00") +
  labs(title = "activity heatmap: steps by hour and day of week") +
  theme_bellabeat()
```

**Insight**: Clear activity peaks during midday hours (12 PM - 7 PM) across all days. Lower activity on weekends compared to weekdays.

**Peak Activity Hours**
```{r eval=FALSE}
ggplot(hourly_activity_summary, aes(x = hour, y = avg_steps)) +
  geom_line(color = "#1976D2", linewidth = 1.2) +
  geom_point(color = "#1976D2", size = 2) +
  geom_area(alpha = 0.2, fill = "#1976D2") +
  labs(title = "average steps by hour of day",
       subtitle = "peak activity hours for notification timing") +
  theme_bellabeat()
```

**Insight**: Optimal times for activity notifications: 12 PM - 2 PM and 5 PM - 7 PM when users are most active.

**Weekday vs Weekend**
```{r eval=FALSE}
ggplot(weekday_weekend_data, aes(x = metric, y = value, fill = period)) +
  geom_col(position = "dodge", alpha = 0.8) +
  labs(title = "weekday vs weekend activity comparison") +
  theme_bellabeat()
```

**Insight**: Users show classic commuter pattern with higher weekday activity, suggesting work-related movement.

### 5.2.3 Correlations

**Correlation Matrix**
```{r eval=FALSE}
ggplot(cor_matrix_long, aes(x = variable, y = variable2, fill = correlation)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#E53935", mid = "white", high = "#43A047") +
  geom_text(aes(label = round(correlation, 2)), size = 2.5) +
  labs(title = "correlation matrix: activity metrics") +
  theme_bellabeat()
```

**Insight**: Strong positive correlations between steps, distance, and calories. Weak negative correlation between sedentary time and calorie burn.

**Steps vs Calories**
```{r eval=FALSE}
ggplot(daily_activities, aes(x = total_steps, y = calories)) +
  geom_point(alpha = 0.3, color = "#1976D2") +
  geom_smooth(method = "lm", color = "#E53935", se = TRUE) +
  labs(title = "steps vs calories burned",
       subtitle = paste0("correlation: ", round(cor_value, 3))) +
  theme_bellabeat()
```

**Insight**: Linear relationship between steps and calories (r > 0.7) validates device accuracy and user behavior patterns.

### 5.2.4 Temporal Trends

**Weekly Activity Trends**
```{r eval=FALSE}
ggplot(weekly_trends, aes(x = week_number, y = avg_steps)) +
  geom_line(color = "#1976D2", linewidth = 1.2) +
  geom_point(color = "#1976D2", size = 3) +
  labs(title = "weekly activity trends",
       subtitle = "average steps per week over 2-month period") +
  theme_bellabeat()
```

**Insight**: Consistent activity levels throughout the study period with no significant drop-off or novelty effect.

**User Retention**
```{r eval=FALSE}
ggplot(weekly_trends, aes(x = as.factor(week_number), y = unique_users)) +
  geom_col(fill = "#43A047", alpha = 0.8) +
  labs(title = "user retention by week") +
  theme_bellabeat()
```

**Insight**: Strong user retention throughout the 2-month period, indicating sustained engagement with tracking.

### 5.2.5 Sleep Visualizations

**Sleep Duration Distribution**
```{r eval=FALSE}
ggplot(sleep_data_plot, aes(x = sleep_hours)) +
  geom_histogram(bins = 30, fill = "#6A1B9A", alpha = 0.7) +
  geom_vline(xintercept = c(7, 9), linetype = "dashed", 
             color = "#E53935", linewidth = 1) +
  annotate("rect", xmin = 7, xmax = 9, ymin = 0, ymax = Inf, 
           alpha = 0.1, fill = "#43A047") +
  labs(title = "sleep duration distribution",
       subtitle = "shaded area shows recommended 7-9 hours") +
  theme_bellabeat()
```

**Insight**: Most users fall within or near the recommended 7-9 hours of sleep, with some outliers on both ends.

**Sleep Efficiency by Activity Level**
```{r eval=FALSE}
ggplot(sleep_by_segment, 
       aes(x = activity_level, y = sleep_efficiency, 
           fill = activity_level)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "sleep efficiency by activity level") +
  theme_bellabeat()
```

**Insight**: Moderate positive relationship between activity level and sleep efficiency, supporting the activity-sleep health connection.

### 5.2.6 Business Insights

**Steps Distribution with 10k Goal**
```{r eval=FALSE}
ggplot(daily_activities, aes(x = total_steps)) +
  geom_histogram(bins = 30, fill = "#1976D2", alpha = 0.7) +
  geom_vline(xintercept = 10000, linetype = "dashed", 
             color = "#E53935", linewidth = 1.2) +
  annotate("text", x = 10000, y = Inf, label = "10k goal") +
  labs(title = "daily steps distribution",
       subtitle = "red line shows 10,000 step goal") +
  theme_bellabeat()
```

**Insight**: Significant portion of users do not consistently meet the 10,000 step goal, presenting an opportunity for goal-based features and encouragement.

---

# 6. Act Phase

## 6.1 Key Insights Summary

Based on comprehensive analysis of FitBit usage data, the following patterns have been identified:

### High Adoption Features
- **Activity Tracking (100%)**: Universal adoption indicates this is the core value proposition
- **Sleep Tracking (80%)**: High adoption suggests users value sleep insights
- Strong engagement with step counting and calorie tracking

### Medium Adoption Features
- **Heart Rate Monitoring (47%)**: Split adoption suggests feature is valued by fitness enthusiasts
- Continuous monitoring appeals to health-conscious segment

### Low Adoption Features
- **Weight Logging (27%)**: Very low adoption indicates:
  - Feature may be too difficult to use
  - Users may not see value in manual logging
  - Opportunity for automated weight tracking integration

### User Behavior Patterns
- **Weekday vs Weekend**: Higher weekday activity suggests commuting patterns
- **Peak Hours**: 12 PM - 7 PM activity peaks indicate lunch breaks and post-work activity
- **Strong Correlations**: Steps and active minutes strongly predict calorie burn
- **Consistent Engagement**: No significant drop-off over 2-month period

## 6.2 Recommendations for Bellabeat

### Strategy 1: Emphasize Core Features

**Objective**: Focus marketing on universally valued features

**Tactics**:
- Highlight activity tracking accuracy and reliability
- Emphasize sleep tracking as premium feature (80% adoption)
- Create comparison content: "Unlike other devices, Bellabeat tracks what matters"
- Position Bellabeat as focusing on the 20% of features that provide 80% of value

**Expected Impact**: Attract users tired of feature bloat in competitor devices

### Strategy 2: Simplify Weight Tracking

**Objective**: Improve adoption of underutilized feature

**Tactics**:
- Integrate with smart scales for automatic weight sync
- Reduce friction in manual logging process
- Add weight trend visualizations and insights
- Connect weight data to activity and sleep for holistic view
- **Alternative**: Consider deprioritizing if integration proves difficult

**Expected Impact**: Increase weight tracking adoption from 27% to 50%+

### Strategy 3: Target Time-of-Day Engagement

**Objective**: Leverage peak activity hours for user engagement

**Tactics**:
- Send motivational notifications during peak hours (12-2 PM, 5-7 PM)
- Offer lunchtime activity challenges
- Create "Evening Wind-Down" feature for 7-9 PM transition to sleep mode
- Avoid off-peak notifications (late night, early morning)

**Expected Impact**: Increase user engagement and app opens by 30%

### Strategy 4: Weekend Warrior Program

**Objective**: Address weekend activity decline

**Tactics**:
- Create "Weekend Challenge" feature encouraging weekend activity
- Offer bonus points/badges for weekend active minutes
- Suggest weekend activities based on weather and location
- Send Friday evening "Plan Your Active Weekend" notifications

**Expected Impact**: Increase weekend activity by 20%, improve weekly consistency

### Strategy 5: Sleep-Activity Connection

**Objective**: Highlight relationship between activity and sleep quality

**Tactics**:
- Create "Active for Better Sleep" educational content
- Show users correlation between their activity and sleep efficiency
- Offer personalized recommendations: "Try 30 more active minutes for better sleep"
- Add "Sleep Readiness Score" based on daily activity

**Expected Impact**: Increase both activity and sleep tracking engagement

### Strategy 6: Segmented Marketing Approach

**Objective**: Target different user segments with tailored messaging

**Segments & Messaging**:

**A. Sedentary Users (<5,000 steps)**
- Focus on simple, achievable goals
- Emphasize health benefits of small increases
- Gentle encouragement, avoid guilt messaging
- "Every step counts" positioning

**B. Active Users (10,000-12,500 steps)**
- Challenge-based features
- Performance optimization
- Social comparison tools
- "Level up your fitness" positioning

**C. Sleep-Focused Users (80% with sleep data)**
- Sleep optimization content
- Bedtime routine features
- Sleep cycle analysis
- "Master your sleep" positioning

**Expected Impact**: 25% increase in conversion rates through personalized messaging

### Strategy 7: Bellabeat App Enhancement

**Objective**: Develop features based on usage patterns

**Priority Features**:
1. **Smart Notifications**: Context-aware reminders during peak hours
2. **Weekly Summary**: Highlight achievements and patterns
3. **Goal Setting**: Personalized step and sleep goals based on user segment
4. **Streak Tracking**: Encourage consistency with visual streak displays
5. **Social Features**: Optional sharing with friends/family for accountability

**Expected Impact**: Increase daily active users by 40%

## 6.3 Product Development Priorities

### Tier 1: Must Have
- Accurate activity tracking (steps, distance, calories)
- Comprehensive sleep tracking with efficiency metrics
- Long battery life (reduce charging friction)
- Water resistance (enable 24/7 wear)
- Comfortable design for extended wear

### Tier 2: Strong Differentiators
- Advanced sleep insights (sleep stages, recommendations)
- Heart rate monitoring for fitness enthusiasts
- Stress tracking integration
- Menstrual cycle tracking (Bellabeat's unique advantage)
- Meditation/mindfulness features

### Tier 3: Nice to Have
- Automatic weight sync (via smart scale integration)
- GPS tracking for outdoor activities
- Music controls
- Payment integration

## 6.4 Success Metrics

Track the following KPIs to measure strategy effectiveness:

### Engagement Metrics
1. **Daily Active Users (DAU)**: Target 30% increase
2. **Feature Adoption Rates**: Sleep (80%→90%), Weight (27%→50%)
3. **Average Session Duration**: Target 15% increase
4. **Notification Open Rate**: Target 40% open rate
5. **Weekend Activity**: Target 20% increase in weekend steps

### Business Metrics
1. **User Acquisition Cost (UAC)**: Monitor campaign efficiency
2. **Customer Lifetime Value (CLV)**: Track long-term engagement
3. **Monthly Active Users (MAU)**: Target steady growth
4. **Churn Rate**: Target <5% monthly churn
5. **Net Promoter Score (NPS)**: Target 50+ score

### Health Impact Metrics
1. **Users Meeting 10k Steps**: Track percentage improvement
2. **Sleep Quality Improvement**: Measure efficiency gains over time
3. **Consistency Streaks**: Track continuous usage periods

## 6.5 Next Steps

### Short-term (0-3 months)
1. Implement smart notification timing based on hourly analysis
2. Launch "Weekend Warrior" challenge program
3. Develop segmented email campaigns for different user types
4. A/B test messaging focusing on sleep vs activity benefits

### Medium-term (3-6 months)
1. Integrate smart scale for automatic weight tracking
2. Develop and launch Bellabeat app enhancements
3. Create comprehensive sleep-activity educational content
4. Launch social features pilot program

### Long-term (6-12 months)
1. Full product line refresh based on findings
2. Expand to new user segments (pregnancy tracking, menopause support)
3. Develop premium subscription tier with advanced insights
4. Build community platform for user engagement and support

---

# 7. Technical Appendix

## 7.1 Data Processing Pipeline

```
Raw Data (data/raw/ - 18 CSV files)
    ↓
Data Cleaning (scripts/01_data_cleaning.R)
    ↓
Cleaned Data (data/cleaned/ - 18 cleaned CSV files)
    ↓
Exploration (scripts/02_exploration.R)
    ↓
Analysis (scripts/03_analysis.R)
    ↓
Processed Data (data/processed/*.csv)
    ↓
Visualization (scripts/04_visualization.R)
    ↓
Final Plots (plots/*.png - 20 visualizations)
```

## 7.2 Key Scripts

### `01_data_cleaning.R`
- Loads 18 raw CSV files
- Standardizes column names
- Validates data quality
- Calculates derived metrics
- Outputs cleaned datasets

### `02_exploration.R`
- Analyzes data coverage
- Calculates feature utilization
- Identifies basic distributions
- Detects outliers
- Generates summary statistics

### `03_analysis.R`
- Performs user segmentation
- Calculates correlations
- Analyzes temporal patterns
- Conducts sleep analysis
- Generates business insights

### `04_visualization.R`
- Creates 20 publication-ready plots
- Implements custom Bellabeat theme
- Generates overview dashboard
- Produces correlation visualizations
- Creates business insight graphics

## 7.3 Code Repository

Complete analysis code available at: [GitHub Repository](https://github.com/yourusername/bellabeat-analysis)

**Repository Structure**:
- `/data/raw/` - Original 18 CSV files
- `/data/cleaned/` - Processed datasets
- `/data/processed/` - Analysis outputs
- `/scripts/` - 4 R analysis scripts
- `/plots/` - 20 generated visualizations
- `/analysis_guide.Rmd` - Comprehensive methodology guide

## 7.4 Reproducibility

To reproduce this analysis:

1. Download FitBit data from [Kaggle](https://www.kaggle.com/datasets/arashnic/fitbit)
2. Place all 18 CSV files in `data/raw/`
3. Run scripts in order:
   ```r
   source("scripts/01_data_cleaning.R")
   source("scripts/02_exploration.R")
   source("scripts/03_analysis.R")
   source("scripts/04_visualization.R")
   ```
4. View outputs in `data/processed/` and `plots/`

**Required R Packages**:
```r
install.packages(c(
  "tidyverse", "dplyr", "ggplot2", "lubridate", 
  "forcats", "vroom", "janitor", "here", "readr", 
  "scales", "broom"
))
```

---

# 8. Conclusion

This comprehensive analysis of FitBit smart device usage data reveals clear patterns in user behavior, feature adoption, and engagement. The findings demonstrate that users value core activity and sleep tracking features (100% and 80% adoption respectively), while more complex features like weight logging see lower adoption (27%).

The analysis identifies five distinct user segments based on activity levels, each requiring tailored marketing approaches. Temporal patterns reveal optimal times for user engagement (12 PM - 7 PM) and highlight opportunities to increase weekend activity.

The recommended seven-strategy approach provides Bellabeat with actionable insights to:
1. Focus marketing on universally valued features
2. Simplify underutilized features like weight tracking
3. Leverage optimal engagement times for notifications
4. Address weekend activity decline
5. Highlight the sleep-activity connection
6. Implement segmented marketing
7. Enhance the Bellabeat app with data-driven features

By implementing these recommendations and tracking defined success metrics, Bellabeat can position itself as a leader in the smart device market, focusing on the features that matter most to women's health and wellness